---
title: "bootstrap"
output: pdf_document
date: "2024-05-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
normal_samples <- rnorm(100, mean = 4, sd = 2)

# Plotting the histogram of normally distributed random numbers
hist(normal_samples, main = "Histogram of Normally Distributed Random Numbers",
     xlab = "Value", ylab = "Frequency", col = "lightblue", border = "black")

# Uniformly distributed random numbers
set.seed(123)  # Setting seed for reproducibility
uniform_samples <- runif(100, min = 0, max = 10)

# Plotting the histogram of uniformly distributed random numbers
hist(uniform_samples, main = "Histogram of Uniformly Distributed Random Numbers",
     xlab = "Value", ylab = "Frequency", col = "lightgreen", border = "black")
```

$P(X \leq z)$。

```{r}
cumulative_prob_at_one <- pnorm(1, mean = 0, sd = 1)
print(cumulative_prob_at_one)


#prob_b <- pnorm(b, mean = mean, sd = sd)
#prob_a <- pnorm(a, mean = mean, sd = sd)
#prob_between_a_and_b <- prob_b - prob_a
#print(prob_between_a_and_b)

quantile_at_95 <- qnorm(0.95, mean = 0, sd = 1)
print(quantile_at_95)
```
### What would be a suitable statistical test for these data and why?

These data are categorical but seriously lacking in independence. Students may have been customers more
than once and, similarly, there are likely to be students who were customers in both trials. The two categories
are therefore not exclusive either.\
• Chi-squared or Fisher’s exact test would be inappropriate\
• Bootstrapping is the only possible alternative to take account of these complications\

##  What are your null and alternative hypotheses? (4 points)

• Null hypothesis: there is no difference between the proportion of students who are satisfied with an
early or late opening time\
• Alternative hypothesis: there is a difference between the proportion of students or students prefer a
late opening time (this would be the equivalent of a one-tailed test)\

```{r}
bootstrap_house_proportion <- function(data, category, num_samples = 1000) {
  # 初始化列表以存储所有重复实验的比例结果
  proportions <- vector("numeric", length = num_samples)
  
  # 进行 num_samples 次有放回的重采样
  for (i in 1:num_samples) {
    sample_data <- sample(data, size = length(data), replace = TRUE)
    category_count <- length(subset(sample_data, sample_data == category))
    proportion <- category_count / length(data)
    proportions[i] <- proportion
  }
  
  # 计算类别数量的 2.5% 和 97.5% 分位数（即置信区间）
  ci <- quantile(proportions, probs = c(0.025, 0.975))
  
  # 将置信区间的下限和上限分别除以样本总数，得到类别比例的置信区间
  lower_bound <- ci[1]
  upper_bound <- ci[2]
  cat("95% Confidence Interval for Proportion of", category, ":", lower_bound, "-", upper_bound, "\n")
  
  return(proportions)
}

data_1 <- c(rep("satisfied", 864), rep("unsatisfied", 714))
data_2 <- c(rep("satisfied", 980), rep("unsatisfied", 473))

# 使用 Bootstrap 函数获取1000次重复实验中 "satisfied" 的比例结果的列表
bootstrap_result_1 <- bootstrap_house_proportion(data_1, "satisfied")
bootstrap_result_2 <- bootstrap_house_proportion(data_2, "satisfied")

# 打印前几个结果查看
print(head(bootstrap_result_1))
print(head(bootstrap_result_2))

result_bootstrap <- data.frame(result_1 = bootstrap_result_1,
                               result_2 = bootstrap_result_2)
result_bootstrap <- result_bootstrap %>%
  pivot_longer(cols = c("result_1","result_2"),
               names_to = "Group",
               values_to = "Result")
result_bootstrap$group <- as.factor(result_bootstrap$Group)
ggplot(data = result_bootstrap,
       mapping = aes(x = Group, y = Result, color = Group))+
  geom_boxplot()
```






# 例子
```{r}
# 定义 Bootstrap 函数
bootstrap_house_proportion <- function(data, category, num_samples = 1000) {
  # 初始化向量以存储每次重采样中指定类别的数量
  category_counts <- vector()
  
  # 进行 num_samples 次有放回的重采样
  for (i in 1:num_samples) {
    sample_data <- sample(data, size = length(data), replace = TRUE)
    category_count <- length(subset(sample_data, sample_data == category))
    category_counts <- c(category_counts, category_count)
  }
  
  # 计算类别数量的 2.5% 和 97.5% 分位数（即置信区间）
  ci <- quantile(category_counts, probs = c(0.025, 0.975))
  
  # 将置信区间的下限和上限分别除以样本总数，得到类别比例的置信区间
  lower_bound <- ci[1] / length(data)
  upper_bound <- ci[2] / length(data)
  
  return(list(mean = mean(category_counts) / length(data), 
              lower = lower_bound, 
              upper = upper_bound))
}

# 示例数据
noncoding <- c(rep("house", 21), rep("other", 292))

# 使用 Bootstrap 函数计算 "house" 的比例置信区间
bootstrap_result <- bootstrap_house_proportion(noncoding, "house")
print(bootstrap_result)
```


