---
title: "CATEGORICAL"
output: pdf_document
date: "2024-05-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
```

## Import and organize the data

```{r}
genotype <- read.csv("genotype.csv")
head(genotype)
```

```{r}
str(genotype)
```

```{r}
cross_table <- table(genotype$sex, genotype$genotype)
print(cross_table)
```

## Describe the data in a useful way

```{r}
ggplot(data = genotype, aes(x = genotype, fill = sex)) +
  geom_bar(position = "dodge") +
  labs(x = "Genotype", y = "Count", fill = "Sex") +
  ggtitle("Genotype Distribution by Sex")
```

### Table 1

|         | $WT$ | $het$ | $mut$ |
|:-------:|:----:|:-----:|:-----:|
| Females |  7   |  26   |   5   |
|  Males  |  10  |  30   |   2   |

## What would you expect under Mendelian inheritance?

In this case, you would expect mice to have males and females as 50/50 and WT, heterozygotes, and mutant mice as 25/50/25.\

### Table 2

|         | $WT$ | $het$ | $mut$ |
|:-------:|:----:|:-----:|:-----:|
| Females |  10  |  20   |  10   |
|  Males  |  10  |  20   |  10   |

#### Choose and justify the appropriate statistical test, state the statistical hypotheses, and carry the test out an appropriate test on whether the mutation affects the survival of mice.

Clearly, it is an expected vs observed distribution. Thus, we need to use the CHI\^2 goodness-of-fit test.\

Assumptions include:\

• The variables must be categorical. – Fits\
• Observations must be independent. – Can assume from the task. Fits.\
• Cells in the contingency table are mutually exclusive.– Fits\
• The expected value of cells should be 5 or greater in at least 80% of cells.– See Table 2. Fits. (may not see Fisher's exact test)

#### State the statistical hypotheses

H0: The data follow the expected distribution (Table 2).\
HA: The data does not follow the expected distribution (Table 2).\

# Simulation

```{r}
popu <- c(rep("WT_M", 5000), rep("het_M", 10000), 
          rep("mut_M", 5000), rep("WT_FM", 5000), 
          rep("het_FM", 10000), rep("mut_FM", 5000))

chi <- function(){
  sam <- sample(popu, 80, replace = FALSE)
  n_WT <- length(which(sam == "WT_M"))
  n_het <- length(which(sam == "het_M"))
  n_mut <- length(which(sam == "mut_M"))
  n_WT_F <- length(which(sam == "WT_FM"))
  n_het_F <- length(which(sam == "het_FM"))
  n_mut_F <- length(which(sam == "mut_FM"))
  X2 <- (n_WT-10)^2/10 + (n_het-20)^2/20 + 
    (n_mut-10)^2/10 + (n_WT_F-10)^2/10 + 
    (n_het_F-20)^2/20 + (n_mut_F-10)^2/10 
  return(X2)
}
chi_sti <- replicate(10000, chi())
Obs <- c(7,20,5,10,30,2)
chi_survey <- sum((7-10)^2/10 + (20-20)^2/20 + (5-10)^2/10 +
                  (10-10)^2/10 + (30-20)^2/20 + (2-10)^2/10)
p.val <- length(which(chi_sti >= chi_survey)) / 10000
plot(density(chi_sti), main = "Simulations of Chi-Square Statistic", xlab = "Chi-Square Value")
abline(v = chi_survey, col = "red") 
print(p.val)
#mock 
chisq.test(Obs, simulate.p.value = TRUE, p = c(0.125, 0.25, 0.125, 0.125, 0.25, 0.125))
```



# Carry test out an appropriate test

As we need to run CHI\^2for goodness-of-fit, we need to provide the expected values. In chisq.test(), we can do that by providing a matrix of expected probabilities under the p argument.\

## Table 3: The expected probabilities

|         | $WT$  | $het$ | $mut$ |
|:-------:|:-----:|:-----:|:-----:|
| Females | 0.125 | 0.25  | 0.125 |
|  Males  | 0.125 | 0.25  | 0.125 |

### Use all the data for chi goodness fit

```{r}
observed <- c(26, 5, 7, 30, 2, 10)  #directly 1-d OK!!
expected <- c(0.25, 0.125, 0.125, 0.25, 0.125, 0.125)
test_result <- chisq.test(observed, p = expected)
print(test_result)
```

### Most easy version

```{r}
tulip <- c(81, 50, 27)
res <- chisq.test(tulip, p = c(1/2, 1/3, 1/6))
res
res$p.value
```

p-value = 0.005324 \< 0.05, we can reject H0 now.\
Thus, the distribution of sex and genotype deviates from the one under Mendelian inheritance law. The mutation did affects the survival of mice.\

Chi-square test of independence:the null hypothesis is that two variables (criteria of classifications defined for the same population) are independent, so the distribution of one of them in no way depends on the distribution of the other.\
Chi-square test of homogeneity:this test is a generalization of the Z/ t test for the difference between two population proportions. It can be used to test the null hypothesis that several population proportions. It can be used to test the null hypothesis that several populations are homogeneous in the sense that categories of a single qualitative variable, or class intervals of a single quantitative variable, have the same distribution in these populations.\
Similarly to the chi-square test of independence, the chi-square test of homogeneity is also based on a contingency table, and the two tests have the same test statistic. However, the two tests differ in terms of the underlying rationale. in the test of independence the expected (joint) frequencies are based on the multiplication rule for independent events, while in the test of homogeneity they are calculated from the pooled sample.\

# Chi-square test of homogeneity

##### Question: Is there a difference between the distribution of the allergic reaction and the preferred seasons?

Assumptions include:\

• The variables must be categorical. – Fits\
• Observations must be independent. – Can assume from the task. Fits.\
• Cells in the contingency table are mutually exclusive.– Fits\
• The expected value of cells should be 5 or greater in at least 80% of cells.– See Table 2. Fits.(may not see Fisher's exact test)\

#### State the statistical hypotheses

H0: The distribution of allergic reactions is the same for the people who preferred different seasons.\
HA: The distribution of allergic reactions is $not$ the same for the people who preferred different seasons.\

## Table 4: The allergic reaction and the preferred seasons

| Category | Spring | Summer | Fall | Winter |
|:--------:|:------:|:------:|:----:|:------:|
|  Severe  |   5    |   1    |  1   |   9    |
|   Mild   |   8    |   5    |  2   |   5    |
| Sporadic |   9    |   8    |  3   |   9    |
|  Never   |   18   |   16   |  12  |   5    |

```{r}
Severe <- data.frame(Spring = 5, Summer = 1, Fall = 1, Winter = 9)
Mild <- data.frame(Spring = 8, Summer = 5, Fall = 2, Winter = 5)
Sporadic <- data.frame(Spring = 9, Summer = 8, Fall = 3, Winter = 9)
Never <- data.frame(Spring = 18, Summer = 16, Fall = 12, Winter = 5)
Two_categories <- rbind(Severe,Mild,Sporadic,Never)
chisq.test(Two_categories)# NOTHING IS IN THE LEFTEST COLOMN

#try fisher.test for better convincing
fisher.test(Two_categories,workspace = 2e7)

Two_categories <- Two_categories %>%
  mutate(Category = c("Severe","Mild","Sporadic","Never"))


Two_categories_long <- Two_categories %>%
  pivot_longer(cols = c(Spring, Summer, Fall, Winter),
               names_to = "Season", values_to = "Count")
# Bubble Plot
ggplot(Two_categories_long, aes(x = Season, y = Category, size = Count, color = Season)) +
  geom_point(alpha = 0.7) +
  labs(title = "Bubble Plot", x = "Season", y = "Category",
       size = "Count") 

```

# Chi-square test of independence

##### Question: Does geneX affect lifespan of mice?

H0: The lifespan is independent on genX H1: The lifespan is dependent on genX

## Table 5:Observed data

|       | WT  | KO  | Total |
|:-----:|:---:|:---:|:-----:|
| Alive |  7  |  2  |   9   |
| Dead  |  3  |  7  |  10   |
| Total | 10  |  9  |  19   |

## Table 6:Expected data

|       | WT  | KO  |
|:-----:|:---:|:---:|
| Alive | 4.7 | 4.3 |
| Dead  | 5.3 | 4.7 |

```{r}
## Fisher (1962, 1970), Criminal convictions of like-sex twins
Convictions <- matrix(c(2, 10, 15, 3), nrow = 2,
	              dimnames =
	       list(c("Dizygotic", "Monozygotic"),
		    c("Convicted", "Not convicted")))
Convictions
fisher.test(Convictions, alternative = "less")
fisher.test(Convictions, conf.int = FALSE)
fisher.test(Convictions, conf.level = 0.95)$conf.int
fisher.test(Convictions, conf.level = 0.99)$conf.int

## A r x c table  Agresti (2002, p. 57) Job Satisfaction
Job <- matrix(c(1,2,1,0, 3,3,6,1, 10,10,14,9, 6,7,12,11), 4, 4,
           dimnames = list(income = c("< 15k", "15-25k", "25-40k", "> 40k"),
                     satisfaction = c("VeryD", "LittleD", "ModerateS", "VeryS")))
fisher.test(Job) # 0.7827
fisher.test(Job, simulate.p.value = TRUE, B = 1e5) # also close to 0.78

```

# Chi-square test of independence

#### input the data into array

## Table 7

#### Hard to drad a three key table

|       |  WT  |   WT   |  KO  |   KO   |
|:-----:|:----:|:------:|:----:|:------:|
|       | Male | Female | Male | Female |
| Alive |  40  |   34   |  20  |   25   |
| Dead  |  9   |   7    |  15  |   20   |

```{r}
mouse_data <- array(c(40, 9, 34, 7, 20, 15, 25, 20), dim = c(2, 2, 2))
dname <- list(status = c("Alive", "Dead"),
sex = c("Male", "Female"),
Genotype = c("WT", "KO"))
dimnames(mouse_data) <- dname
mouse_data#Divide into two smaller table

##Visulize two photos
mouse_data_WT <- data.frame(Status = c("Alive","Dead"),
                            Male = c(40, 9),
                            Female = c(34, 7))
mouse_data_WT_long <- mouse_data_WT %>%
  pivot_longer(cols = c(Male,Female),
               names_to = "Sex", values_to = "Count")
print(mouse_data_WT_long)
ggplot(mouse_data_WT_long, aes(x = Sex, y = Status, size = Count, color = Sex)) +
  geom_point(alpha = 0.7) +
  labs(title = "Bubble Plot of WT", x = "Sex", y = "Status",
       size = "Count")

## 3 way how to convert visulization

mouse_table <- as.table(mouse_data)
mouse_df <- as.data.frame(mouse_table)
colnames(mouse_df) <- c("status", "sex", "Genotype", "value")
print(mouse_df)

ggplot(mouse_df, aes(x = Genotype, y = value, fill = status)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~sex) +
  labs(x = "Genotype", y = "Count", fill = "Status") +
  theme_minimal()
```

Apply the chi-square test. Note: You cannot use an 3-dimensional object in the chisq.test(). Convert the
array to a table object and find the chi-square test result in the summary of the object. Your result should
match the one in the lecture.\
```{r}
mouse_data <- as.table(mouse_data)
summary(mouse_data)
```

In this analysis, we utilized a three-way chi-square test to assess the relationship between genotype (WT and KO), sex (Male and Female), and status (Alive and Dead).\

H0: There is no interaction among the three factors, implying that genotype, and sex do not affect survival status.\
HA: There $is$ no interaction among the three factors, implying that genotype, and sex do not affect survival status, indicating that at least one factor influences survival rate.\

By organizing the data into a three-dimensional table, we conducted a three-way chi-square test to compare observed values with expected values, determining whether there is an interaction among genotype, sex, and status. The chi-square test results include the chi-square statistic and its corresponding p-value. If the p-value is less than a pre-set significance level (typically 0.05), we reject the null hypothesis, suggesting that there is an interaction among the factors.\

Through a statistical summary, we obtained the results of the three-way chi-square test, allowing us to determine whether we reject the null hypothesis and whether genotype, sex, and status have an interaction effect on survival rate.\
