---
title: "Linear_model"
output: pdf_document
date: "2024-05-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
```

## Using diamonds as example
```{r}
head(diamonds)
str(diamonds)
anyNA(diamonds)
diamonds_no_na <- diamonds[complete.cases(diamonds), ]
anyDuplicated(diamonds)
diamonds_unique <- diamonds[!duplicated(diamonds), ]
```

## Select data for lm
We want know the linear relationship in carat and price and is it influenced by cut.\
So cut = ideal and fair selected
```{r}
data_ideal <- subset(diamonds, cut == "Ideal")[ , c("carat", "price")]
data_fair <- subset(diamonds, cut == "Fair")[ , c("carat", "price")]
head(data_fair)
```

Firstly caculating the correlation of carat and price in both cut quality\

```{r}
correlation_ideal <- cor.test(data_ideal$carat,data_ideal$price, use = "complete.obs")
correlation_fair <- cor.test(data_fair$carat,data_fair$price, use = "complete.obs")
print(correlation_ideal)
print(correlation_fair)
```

Both value greater than 0.85 and it can said that in two different cuts, there is an linear relationship exists.\

## Model building and Evaluation
```{r}
model_ideal <- lm(data_ideal$carat ~ data_ideal$price)
model_fair <- lm(data_fair$carat ~ data_fair$price)
```

### Evaluate

```{r}
#For ideal
par(mfrow = c(1,3))
hist(residuals((model_ideal),breaks = 5, col = "grey"),
     main = "Histogram od the residuals", xlab = "residuals", cex = 0.6)
plot(model_ideal, which= c(1,2), cex = 0.6)
```

```{r}
#For fair
par(mfrow = c(1,3))
hist(residuals((model_fair),breaks = 5, col = "grey"),
     main = "Histogram od the residuals", xlab = "residuals", cex = 0.6)
plot(model_fair, which= c(1,2), cex = 0.6)
```

Assume it is good enough\

## Visulizaton

#### Only one group
```{r}
library(ggplot2)
p <- ggplot(data_ideal, aes(x = carat, y = price)) 
p <- p + geom_point()
p <- p + geom_smooth(method = "lm", se = FALSE)
print(p)
```

#### Two group need in one data frame

```{r}
data_selected <- subset(diamonds, cut == c("Ideal","Fair"))[ , c("carat", "price","cut")]
p <- ggplot(data_selected, aes(x = carat, y = price, color = cut))
p <- p + geom_point()
p <- p + geom_smooth(method = "lm", se = FALSE)
print(p)
```

## Calculate Z score

Next, z score is calculated to determine whether the slop of two fitted lines are significantly differently.\
H0: There is no difference between the two correlation coefficients.\
H1: There is significant difference between the two correlation coefficients.\

```{r}
# Extract regression coefficients and standard errors
beta1 <- summary(model_ideal)$coefficients["data_ideal$price", "Estimate"]
se_beta1 <- summary(model_ideal)$coefficients["data_ideal$price", "Std. Error"]
beta2 <- summary(model_fair)$coefficients["data_fair$price", "Estimate"]
se_beta2 <- summary(model_fair)$coefficients["data_fair$price", "Std. Error"]
# Calculate z-score
z <- (beta1 - beta2) / sqrt(se_beta1^2 + se_beta2^2)
print(z) # badly
```
Here, **z is greater than 1.96 (significance level 0.05), indicating a significant difference between the two correlation coefficients, so we reject H0.** This suggests that there is a gender difference in the impact of drug addiction, with males being more affected and having a higher growth rate.